---
title: "Portfolio for Computational Musicology"
author: "Daniël Drucker"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(spotifyr)
library(plotly)
library(compmus)
library(cowplot)
```

```{r include=FALSE}
playlist_beatles <- get_playlist_audio_features("", "18HKG7EtGQyvhL7t5lFXw6") |>
  add_audio_analysis()
playlist_nirvana <- get_playlist_audio_features("", "0G935Lqu0vrEVoPNZNFrqP") |>
  add_audio_analysis()
playlist_coldplay <- get_playlist_audio_features("", "1j08FQSS1hD36ldvdVXP3W") |>
  add_audio_analysis()

playlist_full <-
  bind_rows(
    playlist_beatles |> mutate(artist_name = "The Beatles"),
    playlist_nirvana |> mutate(artist_name = "Nirvana"),
    playlist_coldplay |> mutate(artist_name = "Coldplay")
  )
```

### Track-Level Analyses: Tempo, Volume and Duration
```{r}
playlist_full |>
  mutate(
    sections =
      map(
        sections,                                   
        summarise_at,
        vars(tempo, loudness, duration),            
        list(section_mean = mean, section_sd = sd)
      )
  ) |>
  unnest(sections) |>
  ggplot(
    aes(
      x = tempo,
      y = tempo_section_sd,
      colour = artist_name,
      alpha = loudness
    )
  ) +
  geom_point(aes(size = duration / 60)) +
  geom_rug() +
  theme_light() +
  ylim(0, 5) +
  labs(
    x = "Mean Tempo (bpm)",
    y = "SD Tempo",
    title = "Variation in Tempo",
    colour = "Band Name",
    size = "Duration (min)",
    alpha = "Volume (dBFS)"
  )
```

***
The plot on the left gives a more detailed look of the tempo graph made earlier. Here we can see that indeed tracks of Nirvana have a higher tempo and that the tracks of Coldplay and The Beatles are roughly equally distributed.

What is now also interesting to see is the standard deviation of the tempo for each track. There now seems to be a difference in tempo between The Beatles and Coldplay. According to the graph, The Beatles have more variation of tempo within their tracks then Coldplay. When we keep looking at the standard deviation of tempo, Nirvana seems to also be higher in this regard in comparison to the other two bands.

### Introduction
For years on end my most listened to genre in my Spotify Wrapped has been rock music. Not surprising, since a large part of my personal playlists contain this genre. When looking at this, there is a really wide range of different time periods of rock music in my playlists. This got me thinking about what the differences between rock music in different time periods might be. The main focus of this portfolio will thus be on this comparison.

At first the plan was to look at rock as a whole for this comparison, but I soon found out that this would make the comparisons way to brought. This is why I narrowed the corpus down to three bands that each was very influential in their time: The Beatles in the 60's, Nirvana in the 80's/90's and Coldplay in the modern era.

The corpus used for this portfolio consists of three playlists, one for each band:

- The Beatles: https://open.spotify.com/playlist/18HKG7EtGQyvhL7t5lFXw6?si=943a7297af884c30

- Nirvana: https://open.spotify.com/playlist/0G935Lqu0vrEVoPNZNFrqP?si=d99122f1223f407e

- Coldplay: https://open.spotify.com/playlist/1j08FQSS1hD36ldvdVXP3W?si=b0ec1bff23c644e1

The tracks of these playlists come from a selection of Spotify for each band. I copied those to my own playlists and removed the live versions.


### What is the difference is tempo?
``` {r}
total_mean <- mean(playlist_full$tempo)
total_max <- max(playlist_full$tempo)
x_start = 0.5
y_start <- total_mean - 15
x_end <- 0.5
y_end <- total_mean


tempo <- playlist_full |>
  ggplot(aes(x = artist_name, y = tempo, fill = artist_name, alpha = 0.8)) +
  geom_boxplot() +
  geom_hline(yintercept = total_mean,
    color = "grey40", 
    linetype=3) +
  annotate(
    "text",
    x = x_start, y = y_start,
    label = "Total\naverage",
    vjust = 1, size = 3, color = "grey40"
  ) +
  annotate(
    "curve",
    x = x_start, y = y_start,
    xend = x_end, yend = y_end,
    arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
    color = "grey40"
  ) +
  scale_y_continuous(     
    limits = c(0, total_max + 10),
    breaks = c(0, 50, 100, 150, 200),
    minor_breaks = NULL
  ) +
  theme_light() +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank()
  ) +
  labs(
    x = "Band",
    y = "Tempo",
    title = "Tempo Distribution per Band"
  )

ggplotly(tempo)
```

***

The plot on the left shows how the bands differ in the tempo of their tracks. What strikes out first is that the tracks of Nirvana overall have a much higher tempo then both other bands. The median tempo of Nirvana is about 6.5% higher. This difference is even biggest in the top 50 percentile. This was expected beforehand, since Nirvana is known for its high energy music.

What was not expected however, is the great similarity of tempo of The Beatles and Coldplay. The boxplots of both bands on the left is near identical, where q1, q2 and the median are almost identical. It was expected beforehand that the difference in tempo would be a bit more noticeable. 

### Chromagram and Cepstrogram

```{r}
outlier <- "11LmqTE2naFULdEP94AUBa"

chroma <-
  get_tidy_audio_analysis(outlier) |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

plot1a <-
  chroma |>
    mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
    compmus_gather_chroma() |>
    ggplot(
      aes(
        x = start + duration / 2,
        width = duration,
        y = pitch_class,
        fill = value
      )
    ) +
    geom_tile() +
    labs(x = "Time (s)",
         y = NULL,
         fill = "Magnitude",
         title = "Chromogram (Heart Shaped Box - Nirvana)") +
    theme_minimal() +
    scale_fill_viridis_c()


chepstro <-
  get_tidy_audio_analysis(outlier) |>
  compmus_align(bars, segments) |>
  select(bars) |>
  unnest(bars) |>
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"
      )
  ) |>
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"
      )
  )

plot1b <-
  chepstro |>
    compmus_gather_timbre() |>
    ggplot(
      aes(
        x = start + duration / 2,
        width = duration,
        y = basis,
        fill = value
      )
    ) +
    geom_tile() +
    labs(
      x = "Time (s)",
      y = NULL,
      fill = "Magnitude",
      title = "Chepstrogram (Heart Shaped Box - Nirvana)") +
    scale_fill_viridis_c() +
    theme_classic()

plot_grid(plot1a, plot1b, ncol = 1)
```

***

### Self-Simularity Matrices
```{r}
max_popular_beatles <- playlist_beatles[which.max(playlist_beatles$track.popularity), ]
max_popular_nirvana <- playlist_nirvana[which.max(playlist_nirvana$track.popularity), ]
max_popular_coldplay <- playlist_coldplay[which.max(playlist_coldplay$track.popularity), ]

max_popular_nirvana_uri <- max_popular_nirvana$track.uri
max_popular_coldplay_uri <- max_popular_coldplay$track.uri

track1 <-
  get_tidy_audio_analysis("4CeeEOM32jQcH3eN9Q2dGj") |>
  compmus_align(bars, segments) |>
  select(bars) |>
  unnest(bars) |>
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"
      )
  ) |>
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"
      )
  )

track2 <-
  get_tidy_audio_analysis("3AJwUDP919kvQ9QcozQPxg") |>
  compmus_align(bars, segments) |>
  select(bars) |>
  unnest(bars) |>
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"
      )
  ) |>
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"
      )
  )

plot2a <-
  track1 |>
  compmus_self_similarity(pitches, "cosine") |>
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Chroma (Smells Like Teen Spirit)")

plot2b <-
  track1 |>
  compmus_self_similarity(timbre, "cosine") |>
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Timbre (Smells Like Teen Spirit)")

plot2c <-
  track2 |>
  compmus_self_similarity(pitches, "cosine") |>
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Chroma (Yellow)")

plot2d <-
  track2 |>
  compmus_self_similarity(timbre, "cosine") |>
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Timbre (Yellow)")

plot_grid(plot2a, plot2b, plot2c, plot2d, ncol = 2)

```

***
