---
title: "Portfolio for Computational Musicology"
author: "Daniël Drucker"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(spotifyr)
library(plotly)
library(compmus)
library(cowplot)
```

```{r include=FALSE}
rock_classic <- get_playlist_audio_features("", "37i9dQZF1DWXRqgorJj26U")
rock_modern <- get_playlist_audio_features("", "37i9dQZF1DX1helbHcrYM1")

rock_full <-
  bind_rows(
    rock_classic |> mutate(time_period = "Classic"),
    rock_modern |> mutate(time_period = "Modern")
  )

rock_full$track.album.release_date <- as.Date(rock_full$track.album.release_date)

rock_full <- rock_full %>%
  mutate(year = lubridate::year(track.album.release_date), .after = playlist_owner_id)
```

``` {r}
rock_full <- rock_full %>%
  filter(!(time_period == "Classic" & year > 1999) &
         !(time_period == "Modern" & year < 2000))
```

### Chromagram and Cepstrogram

```{r}
outlier <- "11LmqTE2naFULdEP94AUBa"

chroma <-
  get_tidy_audio_analysis(outlier) |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

plot1a <-
  chroma |>
    mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
    compmus_gather_chroma() |> 
    ggplot(
      aes(
        x = start + duration / 2,
        width = duration,
        y = pitch_class,
        fill = value
      )
    ) +
    geom_tile() +
    labs(x = "Time (s)", 
         y = NULL, 
         fill = "Magnitude",
         title = "Chromogram (Heart Shaped Box - Nirvana)") +
    theme_minimal() +
    scale_fill_viridis_c()


chepstro <-
  get_tidy_audio_analysis(outlier) |> 
  compmus_align(bars, segments) |>                     
  select(bars) |>                                      
  unnest(bars) |>                                      
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"             
      )
  ) |>
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"            
      )
  )

plot1b <-
  chepstro |>
    compmus_gather_timbre() |>
    ggplot(
      aes(
        x = start + duration / 2,
        width = duration,
        y = basis,
        fill = value
      )
    ) +
    geom_tile() +
    labs(
      x = "Time (s)", 
      y = NULL, 
      fill = "Magnitude",
      title = "Chepstrogram (Heart Shaped Box - Nirvana)") +
    scale_fill_viridis_c() +                              
    theme_classic()

plot_grid(plot1a, plot1b, ncol = 1)
```

***

A clear outlier in the tempo plot was "Heart-Shaped Box" by Nirvana. It has the highest tempo of all the classical songs. The chromagram uses Euclidian distance. The clearest plot for the Chestogram was given by the bars alignment, euclidian distance and root mean square summarization. 

### Self-simularity matrices
```{r}
# Select the row with the maximum value for 'modern'
max_popular_modern <- subset(rock_full, time_period == 'Modern')[which.max(subset(rock_full, time_period == 'Modern')$track.popularity), ]

# Select the row with the maximum value for 'classic'
max_popular_classic <- subset(rock_full, time_period == 'Classic')[which.max(subset(rock_full, time_period == 'Classic')$track.popularity), ]

max_popular_modern_id <- max_popular_modern$track.uri
max_popular_classic_id <- max_popular_classic$track.uri

track1 <-
  get_tidy_audio_analysis("1JSTJqkT5qHq8MDJnJbRE1") |> 
  compmus_align(bars, segments) |>                     
  select(bars) |>                                      
  unnest(bars) |>                                      
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"             
      )
  ) |>
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"            
      )
  )  

track2 <- 
  get_tidy_audio_analysis("4SqWKzw0CbA05TGszDgMlc") |> 
  compmus_align(bars, segments) |>                     
  select(bars) |>                                      
  unnest(bars) |>                                      
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"             
      )
  ) |>
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"            
      )
  )
  
plot2a <- 
  track1 |>
  compmus_self_similarity(pitches, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Chroma (Every Breath I Take)")

plot2b <- 
  track1 |>
  compmus_self_similarity(timbre, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Timbre (Every Breath I Take)")

plot2c <- 
  track2 |>
  compmus_self_similarity(pitches, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Chroma (I Love You So)")

plot2d <- 
  track2 |>
  compmus_self_similarity(timbre, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Timbre (I Love You So)")

plot_grid(plot2a, plot2b, plot2c, plot2d, ncol = 2)
```


***
Self-simularity matrices of the most popular track in each category.

Modern: `r max_popular_modern$track.name` - `r max_popular_modern$track.artists.name`

Classic: `r max_popular_classic$track.name` - `r max_popular_classic$track.artists.name`

### Introduction

Fot years on end my most listened to genre in my Spotify Wrapped has been rock music. Not surprising, since a large part of my personal playlists contain this genre. When looking at this, there is a really wide range of different time periods of rock music in my playlists. This got me thinking about what the differences between rock music in different time periods might be. The main focus of this portfolio will thus be on this comparison.

For this research, I will first focus on the main differences between Classic Rock and Modern Rock. For this, I will utilise two pre-made playlists by Spotify:

- Classical Rock: https://open.spotify.com/playlist/37i9dQZF1DWXRqgorJj26U?si=61e74e1c835f4df3

- Modern Rock: https://open.spotify.com/playlist/37i9dQZF1DX1helbHcrYM1?si=fcf01e3fa0594af3

At later stages of the project, we can more deeply zoom in on different time periods or subgenres.

There are two potential weaknesses of the corpus used for this research. The first of these is that these playlists do not take different subgenres into account. It is for example known that in classical rock there is a difference between geographic regions. The difference of these subgenres can have an impact on the overall research. I do however not have the knowledge to make these distinctions, so I will just look at the big picture. The second weakness is that the corpus used is made by Spotify, which tends to put recommendations for the user in its playlists. This means the corpus might be a reflection of my own musical taste.

``` {r}
count_year <- rock_full %>%
  group_by(year, time_period) %>%
  summarise(count = n())

years <- ggplot(count_year, aes(x = year, y = count, fill = time_period)) +
  geom_bar(stat = "identity") +
  theme_light() + 
  theme(
    panel.grid.major.x = element_blank()
  ) +
  labs(
    x = "Count",
    y = "Year",
    title = "Distibution of year per playlist",
    fill = "Playlist"
  )

ggplotly(years) %>%
  layout(height = 400, width = 650)
```

### Classic Rock has a slower tempo than Modern Rock

```{r}
total_mean <- mean(rock_full$tempo)
total_max <- max(rock_full$tempo)
x_start = 0.5
y_start <- total_mean - 15
x_end <- 0.5
y_end <- total_mean


tempo <- rock_full |>
  ggplot(aes(x = time_period, y = tempo, fill = time_period, alpha = 0.8)) +
  geom_boxplot() +
  geom_hline(yintercept = total_mean,
    color = "grey40", 
    linetype=3) +
  annotate(
    "text",
    x = x_start, y = y_start,
    label = "Total\naverage",
    vjust = 1, size = 3, color = "grey40"
  ) +
  annotate(
    "curve",
    x = x_start, y = y_start,
    xend = x_end, yend = y_end,
    arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
    color = "grey40"
  ) +
  scale_y_continuous(     
    limits = c(0, total_max + 10),
    breaks = c(0, 50, 100, 150, 200),
    minor_breaks = NULL
  ) +
  theme_light() +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank()
  ) +
  labs(
    x = "Location",
    y = "Tempo",
    title = "Tempo Distribution per Location"
  )

ggplotly(tempo)
```

*** 

Modern Rock Music does seem to have a noticeably higher tempo than Classical Rock. One of the reasons of this can be the overall faster pace of life in the modern world, which might be reflected in music.

The difference, however, is not that big. The median tempo of Modern Rock is only about 3.4% higher than Classical Rock. The main difference seems to be in the top 50 percentile, where Modern Rock has a significantly higher tempo than its classical counterpart. Another interesting observation is the fact that Modern Rock does have a bigger variety.


### Test 2
